---
title: "Creative Assignment 4"
author: "Jocelyn Tsai"
date: "10/2/2020"
output: html_document
---
```{r, message=FALSE}
library(osmdata)
library(opentripplanner)
library(tidyverse)
library(sf)
library(ggthemes)
library(ggspatial)
```

For this assignment, I am interested in looking at the walksheds and drivesheds to Emergency Shelters in Cambridge. It is important for Cambridge residents to be able to quickly access these shelters during emergency situations.

Emergency Shelters data from: 
https://www.cambridgema.gov/GIS/gisdatadictionary/Public_Safety/PUBLICSAFETY_EmergencyShelters

Reading the shapefile that was saved to my computer: 

```{r}
emergency_shelters <- st_read(
  "data_downloads/PUBLICSAFETY_EmergencyShelters.shp/PUBLICSAFETY_EmergencyShelters.shp")
```

The opq function creates a call to the open street map server to download street network data. bbox (bounding box), is like the search bar on google maps (typing in general area name is fine, it doesn't have to be exact). osmdata_xml will create a file with the information about the street network, and it will save it to the default directory created.

```{r}
opq(bbox = 'Cambridge MA USA') %>%
  add_osm_feature(key = 'highway') %>%
  osmdata_xml(file = 'OTP/graphs/default/cambridge_streets.osm')
```

I am transforming my spatial data of street lines onto the NAD83(HARN) / Massachusetts Mainland projected coordinate system from https://spatialreference.org/ref/epsg/2805/ .
The add_osm_feature when 'highway' is specified takes all the roads in Open Street Map. I then used osmdata_sf to also download data in the sf simple features format so I can plot it. 

```{r}
MA_mainland <- "+proj=lcc +lat_1=42.68333333333333 +lat_2=41.71666666666667 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"

cambridge_street_features <- opq(bbox = 'Cambridge MA USA') %>%
  add_osm_feature(key = 'highway') %>%
  osmdata_sf()

cambridge_streets <- cambridge_street_features$osm_lines %>%
  st_transform(crs = MA_mainland)
```

I then plotted the streets in Cambridge to see what's included:

```{r}
ggplot(cambridge_streets) +
  geom_sf() 
```

Download a Java utility called otp.jar and save it to the OPT folder (only need to run this code chunk once):

```{r}
path_otp <- otp_dl_jar("OTP")
```

Build a graph representing street and transit networks:

```{r, message=FALSE, results='hide'}
path_data <- file.path(getwd(), "OTP")
path_otp <- paste(path_data, "otp.jar",sep = "/")

otp_build_graph(otp = path_otp, dir = path_data, memory = 1024) 
```

Setting up Open Trip Planner (OTP application opens in the web browser):

```{r, message=FALSE, results='hide'}
otp_setup(otp = path_otp, dir = path_data, memory =1024)
```

Connect to Open Trip Planner:

```{r}
otpcon <- otp_connect()
```



Create isochrones 

I am creating isochrones for areas within an 8 minute walk and an 8 minute drive: 

```{r}
iso_8min_walk <- 
  otp_isochrone(otpcon = otpcon, fromPlace = emergency_shelters, 
                mode = "WALK", cutoffSec = 480) %>%
  st_transform(crs = MA_mainland) %>%
  mutate(mode = "walk")

iso_8min_drive <- 
  otp_isochrone(otpcon = otpcon, fromPlace = emergency_shelters, 
                mode = "CAR", cutoffSec = 480) %>%
  st_transform(crs = MA_mainland) %>%
  mutate(mode = "drive")

iso_all_modes <- rbind(iso_8min_walk, iso_8min_drive)

otp_stop()
```



