---
title: "Creative Assignment 1"
author: "Jocelyn Tsai"
date: "9/12/2020"
output: html_document
---
```{r, message=FALSE}
library(ggplot2)
library(tidyverse)
library(tidycensus)
```

I decided to use data from the 2010 Census for this assignment. I first loaded all the available variables in the census to see which variables I am interested in picking.
```{r}
Census2010_variables<- load_variables(2010, "sf1", cache = TRUE)

View(Census2010_variables)
```

I ended up choosing the same variables as Professor Voulgaris's written tutorial because I wanted to spend more time on experimenting with ggplot. I used census tract data from Alameda County in California, from the 2010 Census. 

The six variables used are:
  1. Whether the census tract is rural or not (categorical): rural
  2. Percentage of household in rented homes (continuous): pct_rental
  3. The majority race/ethnicity (categorical): majority_race_eth
  4. Average household size (continuous): avg_hh_size
  5. Total tract population (continuous): tot_pop
  6. Tract residents' median age (continuous): med_age

```{r, message=FALSE}
vars <- c(tot_hh = "H001001",
         tot_rural_hh = "H002005",
         tot_rent_hh = "H004004",
         avg_hh_size = "H012001",
         tot_pop = "P003001",
         nh_white_pop = "P005003",
         nh_black_pop = "P005004",
         nh_asian_pop = "P005006",
         hispanic_pop = "P005010",
         med_age = "P013001")

Alameda_tracts <- get_decennial(geography = "tract", variables = vars,
                           state = "CA", county = "Alameda",
                           output = "wide") %>%
  
  mutate(rural = (tot_rural_hh == tot_hh) ,
         pct_rental = tot_rent_hh / tot_hh,
         majority_race_eth = case_when(
           nh_white_pop / tot_pop > 0.5 ~ "nh_white",
           nh_black_pop / tot_pop > 0.5 ~ "nh_black",
           nh_asian_pop / tot_pop > 0.5 ~ "nh_asian",
           hispanic_pop / tot_pop > 0.5 ~ "hispanic",
           (nh_white_pop + nh_black_pop + nh_asian_pop + hispanic_pop) /
             tot_pop < 0.5 ~ "other",
           TRUE ~ "None")) %>%
  
  filter(tot_pop > 0, tot_hh > 0) %>%
  
  select(GEOID, rural, majority_race_eth, 
         pct_rental, avg_hh_size, tot_pop, med_age) 
```

```{r}
ggplot(Alameda_tracts, 
       aes(x = pct_rental,
           y = med_age)) +
  geom_point() +
  scale_x_continuous(name = "Percent of Households in Rented Homes",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  scale_y_continuous(name = "Tract Residents' Median Age",
                     breaks = seq(10, 100, by = 10)) +
  stat_smooth(method = "lm")
  theme_bw()+
  theme(legend.position = "none")
```
```{r}
ggplot(Alameda_tracts, 
       aes(x = majority_race_eth,
           y = avg_hh_size,
           color = majority_race_eth)) +
  geom_point(position = "jitter",
             size = 0.7,
             alpha = 0.8) +
  scale_x_discrete(name = "Majority Race/Ethnicity",
                       labels = c("Hispanic",
                                  "Non-Hispanic Asian",
                                  "Non-Hispanic white",
                                  "No majority")) +
  scale_y_continuous(name = "Average Household Size") +
  theme_bw() +
  theme(legend.position = "none")
```
```{r}
ggplot(Alameda_tracts, 
       aes(x = majority_race_eth,
           y = pct_rental,
           color = majority_race_eth)) +
  geom_boxplot() +
  scale_x_discrete(name = "Majority race/ethnicity",
                       labels = c("Hispanic",
                                  "Non-Hispanic Asian",
                                  "Non-Hispanic white",
                                  "No majority")) +
  scale_y_continuous(name = "Percent of Households in Rented Homes",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  theme_bw() +
  theme(legend.position = "none")
```

type \n in between legend titles to switch lines
problem with this graph is that I can't tell if the darker black circles from a larger Average Household Size or just because the circles are overlapping, which makes it darker.

```{r}
ggplot(Alameda_tracts, 
       aes(x = med_age,
           y = tot_pop,
           size = pct_rental,
           alpha = avg_hh_size)) +
  geom_point() +
  scale_x_continuous(name = "Tract Median Age",
                     breaks = seq(10, 80, by = 10)) +
  scale_y_continuous(name = "Total Tract Population") +
  scale_size_continuous(name = "Percent Households\nin Rentals",
                    breaks = seq(0, 1, by = 0.1),
                    labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  scale_alpha_continuous(name = "Average Tract\nHousehold Size") +
  theme_bw() +
  theme(legend.key.size = unit(0.05, "cm"))
```

I found out that there's no rural tracts in Alameda County because all the fill in the bar chart is the same color. 

```{r}
ggplot(Alameda_tracts, 
       aes(y = majority_race_eth, fill = rural)) +
  geom_bar() +
  scale_y_discrete(name = "Majority Race/Ethnicity",
                       labels = c("Hispanic",
                                  "Non-Hispanic\nAsian",
                                  "Non-Hispanic\nwhite",
                                  "No majority")) +
  scale_x_continuous(name = "Number of tracts") +
  scale_fill_discrete(name = "Urban or Rural Tracts",
                        labels = c("Urban tracts",
                                   "Rural tracts")) +
  theme_bw() 
```

```{r}
ggplot(Alameda_tracts, 
       aes(x = majority_race_eth,
           y = avg_hh_size,
           color = majority_race_eth)) +
  geom_violin() +
  scale_x_discrete(name = "Majority Race/Ethnicity",
                       labels = c("Hispanic",
                                  "Non-Hispanic Asian",
                                  "Non-Hispanic white",
                                  "No majority")) +
  scale_y_continuous(name = "Average Household Size") +
  theme_bw() +
  theme(legend.position = "none")
```

```{r}
ggplot(Alameda_tracts, 
       aes(x = tot_pop,
           y = majority_race_eth,    
           color = majority_race_eth)) +
  geom_violin() +
  geom_point(position = "jitter",
             alpha = 0.7,
             size = 0.7) +
  scale_x_continuous(name = "Total Population within tract") +
  scale_y_discrete(name = "Majority race/ethnicity",
                       labels = c("Hispanic",
                                  "Non-Hispanic Asian",
                                  "Non-Hispanic white",
                                  "No majority")) +
  theme_bw() +
  theme(legend.position = "none")
```

```{r}
ggplot(Alameda_tracts, 
       aes(y = rural,
           x = med_age)) +
  geom_point() +
  scale_y_discrete(name = "Urban or Rural Tracts",
                       labels = c("Urban Tracts",
                                  "Rural Tracts")) +
  scale_x_continuous(name = "Tract Median age",
                     breaks = seq(10, 100, by = 5)) +
  theme_bw() 
```

```{r}
ggplot(Alameda_tracts, 
       aes(x = med_age,
           y = pct_rental)) +
  geom_point(alpha = 0.5, size = 0.5) +
  stat_smooth(color = "purple", linetype = 1, size = 0.8) +
  scale_x_continuous(name = "Tract Median Age",
                     labels = c("100 or 0", "10", "20", "30", "40", "50", 
                                "60", "70", "80", "90", "100"),
                     breaks = seq(0, 100, by = 10)) +
  scale_y_continuous(name = "Percent Households in Rentals",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  coord_polar(theta = "x") +
  theme_bw() 
```

