---
title: "Creative Assignment 7"
author: "Jocelyn Tsai"
date: "11/6/2020"
output: 
  html_document:
    theme: yeti
    highlight: haddock
    toc: true
    toc_float: true
---
```{r, message=FALSE}
library(tidyverse)
library(sf)
library(leaflet)
library(htmltools)
library(htmlwidgets)
library(raster)
library(gstat)
library(spatial)
```

## Dataset

I downloaded social vulnerability data (part for Climate Ready Boston) census tracts within the City of Boston. Data source: https://bostonopendata-boston.opendata.arcgis.com/datasets/34f2c48b670d4b43a617b1540f20efe3_0?geometry=-71.551%2C42.227%2C-70.563%2C42.405

```{r}
social_vulnerability <- st_read(
  "data_downloads/Climate_Ready_Boston_Social_Vulnerability-shp/3aeae140-8174-4d77-8c0e-de3ef0ce4b672020330-1-1rr22uq.veze.shp")
```

I will be looking at percentage of people with disability in each census tract within Boston since they are most vulnerable during disaster scenarios. I added a column for total people with disabilities/ total tract population to get the percentage.

```{r}
social_vulnerability <- social_vulnerability %>%
  mutate(ppl_disability= (social_vulnerability$TotDis/social_vulnerability$POP100_RE))
```

Omit rows with N/A values
```{r}
social_vulnerability_1 <- na.omit(social_vulnerability)
```


## Map 1: chloropleth map

```{r, message=FALSE, results='hide'}
colorbins <- colorBin(palette = "Reds",
                   domain = social_vulnerability_1$ppl_disability,
                   na.color = "#00000000",
                   bins = 9,
                   pretty = FALSE)

Map_1 <- leaflet(social_vulnerability_1, 
                 options=leafletOptions(minZoom = 10, maxZoom = 16)) %>%
  addProviderTiles("Esri.WorldGrayCanvas") %>%
  addPolygons(fillColor = colorbins(social_vulnerability_1$ppl_disability), stroke = FALSE, smoothFactor=0.2, fillOpacity = 1,
              highlightOptions = highlightOptions(fillColor = "purple", fillOpacity = 0.5),
              label = social_vulnerability_1$Name,
              popup = paste("Neighborhood:", social_vulnerability_1$Name, "<br/>",
                            "Tract population count:", social_vulnerability_1$POP100_RE, "<br/>",
                            "Tract population with disability:", social_vulnerability_1$TotDis, "<br/>",
                            "Percentage of population with disability:", format(round((social_vulnerability_1$ppl_disability)*100,digits =2)),"%")) %>%
  
    addControl('<a href="https://bostonopendata-boston.opendata.arcgis.com/datasets/34f2c48b670d4b43a617b1540f20efe3_0?geometry=-71.551%2C42.227%2C-70.563%2C42.405&page=9">Data source</a>',
             position = "bottomleft") %>%
   addControl("Boston People with Disabilities", position = "topright") %>%

  setMaxBounds( lng1 = -71.16,
                lat1 = 42.27,
                lng2 = -70.89,
                lat2 = 42.39) %>%
    addLegend("bottomright",
            pal= colorbins,
            values = social_vulnerability_1$ppl_disability,
            opacity = 1,
            title = "Percentage of population with disability")
  
Map_1
```


## Map 2: Points map using area centroids


```{r, message=FALSE, results='hide'}
MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs "

WGS84 <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"

nhood_points <- st_centroid(
  st_transform(social_vulnerability_1, crs = MA_state_plane)) %>%
  st_transform(WGS84)

leaflet(nhood_points, 
                 options=leafletOptions(minZoom = 10, maxZoom = 16)) %>%
  addProviderTiles("Esri.WorldGrayCanvas") %>%
  
  addCircles(label = social_vulnerability_1$Name,
             fillColor = colorbins(social_vulnerability_1$ppl_disability),
             stroke = FALSE, 
             radius = 150, 
             fillOpacity = 1) %>% 
    addControl('<a href="https://bostonopendata-boston.opendata.arcgis.com/datasets/34f2c48b670d4b43a617b1540f20efe3_0?geometry=-71.551%2C42.227%2C-70.563%2C42.405&page=9">Data source</a>',
             position = "bottomleft") %>%
   addControl("Boston People with Disabilities", position = "topright") %>%

  setMaxBounds( lng1 = -71.16,
                lat1 = 42.27,
                lng2 = -70.89,
                lat2 = 42.39) %>%
    addLegend("bottomright",
            pal= colorbins,
            values = social_vulnerability_1$ppl_disability,
            opacity = 1,
            title = "Percentage of population with disability")

```

 
## Map 3: Interpolation map 

Raster interpolation can make guesses for areas in between points in Map 2. First convert polygons and points to state plane coordinates so distance measurements used in the raster interpolation will be accurate. 

note: "The raster functions, including for spatial interpolation, work with an older data format (Spatial* - or sp) that what we’ve been working with (Simple Features - or sf). That’s fine because it’s pretty easy to convert back and forth" 

```{r, message=FALSE, results='hide'}
nhood_pts_sp <- nhood_points %>%
  st_transform(MA_state_plane) %>%
  as_Spatial()

nhood_poly_sp <- social_vulnerability_1 %>%
  st_transform(MA_state_plane) %>%
  as_Spatial()
```

Create an empty raster that covers the area of Boston (the extent of the census tracts I am mapping) with a resolution of 10 meters (each cell represents 10 meters).

```{r, message=FALSE, results='hide'}
boston_raster <- raster(nhood_poly_sp, res=10)
```

Now use the inverse-distance weighting method to show variation in percent of people with disability in different census tracts. gstat() and interpolate() do the raster interpolation. mask() function clips the raster layer to the extent of Boston census tracts. 

(when I interpolated, it could not run because there was an error message that it's missing values, so I went back to the top and and omitted rows with N/A value)

```{r, message=FALSE, results='hide'}
gs <- gstat(formula=ppl_disability~1, locations=nhood_pts_sp)
idw_interp <- interpolate(boston_raster, gs)

idw_interp_clip <- mask(idw_interp, nhood_poly_sp)
```

```{r, message=FALSE, results='hide'}
leaflet(nhood_points) %>%
  addProviderTiles("Esri.WorldGrayCanvas") %>%
  addRasterImage(idw_interp_clip, colors = colorbins(social_vulnerability_1$ppl_disability), opacity = 0.8) %>% 

    addControl('<a href="https://bostonopendata-boston.opendata.arcgis.com/datasets/34f2c48b670d4b43a617b1540f20efe3_0?geometry=-71.551%2C42.227%2C-70.563%2C42.405&page=9">Data source</a>',
             position = "bottomleft") %>%
   addControl("Boston People with Disabilities", position = "topright") %>%

  setMaxBounds( lng1 = -71.16,
                lat1 = 42.27,
                lng2 = -70.89,
                lat2 = 42.39) %>%
    addLegend("bottomright",
            pal= colorbins,
            values = social_vulnerability_1$ppl_disability,
            opacity = 1,
            title = "Percentage of population with disability")
```

## Discussion 

Out of the 3 maps (Map 1: Chloropleth map, Map 2: Points map, Map 3: Interpolation map), I think the points map representing data at the middle of the tract is the most appropriate, informative, and the best map out of the 3 because the total population of people with disability is aggregated by individual census tracks. Points map also doesn't show sudden changes in percentage of population with disabilities across census tract boundaries unlike the Chloropleth map. 

To me, the interpolation map looks the most interesting but does not make much sense because spatially it seems like areas towards the border of the census tracts have either higher or lower percentage of people with disabilities comparing to the center point of the tract when in fact we do not know the spatial distribution of where people with disabilities are located within any census tracts. I also do not understand why there are ripple-like colors on my interpolation map representing variation in percentages. 

